Ligeti, Stockhausen and Koenig:\\Aesthetics of Electronic Music

### 2015-09-26 manuscript.tex
Modify some concenpts titles. Get ready to comparizon diagram
Add concepts categorization diagram
Complete comparison of ideas

@@

\subsubsection{Lower level building blocks for electronic sound}
\label{subs:stockhausen:sound}

@@

\subsubsection{Sounds of limited vs. unlimited lengths}
\label{subs:koenig:lengths}

@@

TODO regarding the first concept by Stockhausen:
Overall, I can not find a common ground between this compository element to those of Ligeti that were mentioned before.
Their attitudes regarding unified perception of time are, most probably, contradicting.

TODO regarding the 2nd concept by stockhausen:
Ligeti also refer to this ideas:

\begin{MyShadequote}
  Modification of timbre and dynamics are obviously very significant but the patterns emerging from them are even more important...
  I should say that the changes and modifications of the overall pattern are the important feature, not the tone colors. (Ligeti in conversation, p. 39).
\end{MyShadequote}

>>>>

\usepackage{graphicx}

@@

\subsubsection{Lower level building blocks}
\label{subs:stockhausen:sound}

@@

\subsubsection{Sounds of unlimited lengths}
\label{subs:koenig:lengths}

@@

\begin{figure}[!htb]
  \centering
  \textbf{Categorization of ideas}\par\medskip
  \includegraphics[width=\linewidth]{graphics/concepts_categorization.pdf}
  \caption{Subjective categorization of ideas to sound, serializm and high level expressions.}
  \label{fig:concepts_categorization}
\end{figure}

Generally, I found that most of the ideas above can be divided easily into three categories --- sound, serial thinking, and higher level expressions --- as shown in Figure \ref{fig:concepts_categorization}.

The most prominent category in the composers writings is, perhaps, sound.
Both Stockhausen and Koenig count several characteristics of electronic sound which differ them from instrumental music.
Ligeti, on the other hand, composes mainly instrumental music but use very similar terminology to describe the sound he is looking for.
I think that Koenig description of non-stationary sounds in electronic music (see subsection \ref{subs:koenig:nonstationary}) may be used as a superset for all of the other concepts presented in this category.
The ``densely woven cobweb'', ``floating, fluctuating patterns'', and ``dirty smudginess'' of Ligeti are very similar to Stockhausen's ``ordering of the succession of pulses'' and Koenig's ``micro-time movement of timbre''.

It is important to note though that Ligeti's interest in the modification of timbre should be understood as part of a larger context.

\begin{MyShadequote}
  Modification of timbre and dynamics are obviously very significant but the patterns emerging from them are even more important...
  I should say that the changes and modifications of the overall pattern are the important feature, not the tone colors. (\cite[p. 39]{varnai}).
\end{MyShadequote}

The same might be applied to Stockhausen and Koenig as well, whereas the larger context is the serial way of thinking the modification of timbre is part of.
Serial aesthetics in electronic music is beyond the scope of the current work, thereby I will refrain from answering that question.

The second category I found is serializm.
As mentioned earlier, I will not deal directly with serializm in the current work.
However, there are some concepts that are rooted in serial way of thinking that can be generalized and understood by themselves.
Concepts like these, despite their problematic source, are not ignored and will be analyzed in the next section.
This category is probably the most foreign one to Ligeti's way of thinking.
Moreover, it seems that a unified model of time, as described by Stockhausen, contradict much of Ligetis' concepts from the next category.

The last category is the one for higher level expressions, and is occupied almost solely by Ligetis' ideas.
These concepts based on expressions and feelings, and are characterized by descriptions that are not musical in their nature.

Two concept are not categorized simply into these three categories.
Noise, for example, as described by Stockhausen, can be understood in every time range;
e.g. both as a property of the sound itself, as well as of rhythm or form.
In that regard, this idea is belong both to sound and to serial thinking, as it emerged from Stockhausen unified model of time.
In the next section I will try to refer to this concepts in different time ranges.

The last concept is the representation of pitch ranges.
It is clearly a serial concept, but part of it, especially the ``transitional'' representation, remind me the ``Mistiness'' to ``clearing-up'' process of Ligeti.
Extract this idea from its serial way of thinking, and it can be seen as a very expressive description of relationships between ``planes'' of sounds and melodies, including intermediate states.
